Estimation of the road traffic sound levels in urban areas based on non-negative matrix factorization techniques

### 2018-01-18 paper/gloaguenNmfTraffic.tex
AC and ML corrections

\documentclass[twocolumn,a4paper,10pt]{article}

@@

The advent of low cost acoustic monitoring devices raises new interesting approaches for improving the monitoring the acoustic quality of urban areas. Assuming that transportation is the major source of annoyance, state of the art approaches consider as input an estimate of the number and the speed of vehicules in major traffic lanes. Follows a prediction procedure that outputs an acoustic pressure level at any location in the modeled area.

Considering as input the acoustic pressure measured in many locations using a sensor grid approach would greatly complement and improve the quality of the predicted pressure values. Among the technical issues that raises this kind of innovative approaches, there is a need to identify which part of the overall acoustic pressure level is due to the road traffic.

In this paper, several techniques based on the non-negative factorization framework are studied in this application scenario. The task being to the best of our knowledge never been considered in the literature, we propose an exprimental protocol to validate the studied approaches that complies with standard reproducible research recommandations.

@@

With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps. These maps depict an estimation of the number of city dwellers exposed to high noise levels. They are used to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}. These maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation techniques. They express $L_ {DEN}$ and $L_N$, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively. However, these maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015}, by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014} \ml{un peu hors propos on améliore rien la dessus : or even by the calculation procedure of the number of inhabitants exposed to noise \cite{king_implementation_2011}. plutot parler du cote sparse des données d'entrées}. In addition, the usual road traffic noise maps are static, aggregating the exposure on the two indicators $L_{DEN}$ and $L_N$, thus ignoring the sound levels evolution throughout the day. Since the creation of road traffic noise maps entails long data collection and calculation times, the use of acoustic measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}. These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would lead to the availability of the long-term evolution of the traffic noise levels. It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time periods.

Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters ...), including the assessment of urban noise levels. DYNAMAP project \cite{dynamap_2016} studied the deployment and feasibility of such installations. It focuses on sensor installations on specific roads at the city scale in Milan and Rome \cite{bellucci_life_2017}. In a similar way, but reduced to few neighborhoods \ml{with a more dense topology ? parler aussi de sonyc}, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{picaut2017characterization} aims to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, through data assimilation techniques.

If sensor networks could improve road traffic noise estimates compared with simulated maps, the issue of the correct estimation of the traffic sound level from acoustic measurements is still unsolved \cite{Mioduszewski}. Indeed, the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes \dots) that overlap. Consequently, the traffic sound level estimation based on measurements is not a trivial task.
Many recent works have focused on the detection or recognition tasks of environmental sounds \cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}. A two-step process is generally followed : describe the audio files with a set of features (Spectrum Gravity Spectrum, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks). A description of these features and classifiers can be found in \cite{cowling_comparison_2003} and their applications can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.

The main issue in the detection or recognition tasks is the overlap of environmental sounds. Although near major roads or ring roads, traffic is predominant, there are many places where it overlaps with other sound sources that contribute significantly to the overall sound levels. To circumvent this issue, Socoro et al. propose to suppress time frames where there is significant overlap by considering an Anomalous Noise Events Detector \cite{socoro_anomalous_2017} which consist in detecting the unwanted sound sources from labeled recordings, \textit{i.e.} that are not related to the traffic component. Those time frames are then discarded in order not to take them into account during the estimation of the traffic sound level.  \ml{arrete :) If the detection of the road traffic noise is good, the detection of these anomalous noise events stay a complicated task \cite{stowell2015detection}.}

An alternative approach that we will foolow in this paper is to consider the blind source separation paradigm to reliably estimate the traffic noise level. That is, separating the contribution of the traffic from the other sources within a polyphonic scene. One major advantage of following such approach is that the estimate is continuously available, making the approach applicable in a wide range of urban areas, even where the traffic noise is relatively low compared to the remaining contributions.\\

ml{Supprimer : One of the first to do so is the Independent Component Analysis \cite{comon_independent_1994}. The principle is to decompose $N$ recorded signals to a sum of $P$ independent sound sources weighted by linear relations. This method is most of all suited for the 'cocktail party' issue where one tries to capture a signal among noise. In an urban environment context, this method is used with acoustic microphone arrays and beamforming \cite{saruwatari2003blind}. These approaches consider a set of microphones and allows the localization \cite{mennitt2010multiple} or the detection \cite{leiba2017large} of sound sources from the phase shift and the distance between the microphones. However, one need a lot of microphones. Spread multiples microphones arrays in cities is then expensive (even with low cost microphones) and is time-consuming for calibration and maintenance.}

A convenient method for monophonic signal decomposition is the Non-negative Matrix Factorization (NMF) technique \cite{lee_learning_1999}. When considering audio as input, it usually consists in approximating the magnitude spectrogram of an audio file by the product of two low rank matrices, one representing the components of interest and the other the contribution at a given time of those components to approximate the input magnitude spectrogram \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}. In the audio processing domain, the NMF has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}. By design, this method deals reasonably well with the overlaping sound sources as soon as the overlap can be resolved on the time/frequency plane. \ml{enlever : For the environmental sounds, the method has been used for the geo-localisation and classification of the sound environment, like in \cite{kumar_audio_2016} where NMF is used to classify the audio files according to the 10 cities where they have been recorded.}

Closer to our application scenario, NMF has been considered by Innami and Kasai \cite{satoshi_innami_nmf-based_2012}. After having performed NMF on simulated audio files, they perform a source separation based on NMF in two steps by 1) separating the sound background from the events and 2) by isolating the events using spectral features using a $k$-means procedure.

\ml{ enlever : A $k$-mean clustering allows the separation of the $k$ sound sources. This approach, compared to a simple MFCC clustering, improves the sound source separation but the authors admit their approach need the assumption that the number of sound sources is known which is unrealistic in practical case.  Furthermore, in order to estimate the traffic sound level $\tilde{L}_{p,traffic}$, their approach might be difficult to be adapted it as traffic component can be both sound background and event.}

We study in this paper different flavors of Non-Negative Matrix Factorization (NMF) where traffic component is considered in its entirely whether it is a sound background or event. We demonstrate the supervised NMF and semi supervised NMF approaches have some interests but fail to give satifactory results for the application at hand. We thus introduce another NMF scheme called thresholded initialized NMF that make good use of prior knowledge about the source of interest, in our case the traffic noise, but also generalizes well to several kinds of urban areas and to traffic to interference ratio (TIR).

To perform the numerical experiments, we consider a corpus of simulated sound scenes created from a built-up sound database composed of a high number of diverse sound samples. The use of simulated sound scenes is mandatory for rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$) which would be difficult to extract from a recording of an urban scene.

\begin{figure*}[t]
\centering
\includegraphics[width=0.7\textwidth]{figures/bloc_diagram_source_separation.pdf}
\caption{Block diagram of the source separation method. \ml{toujours citer une figure}}
\label{fig:diagram}
\end{figure*}

The remaining of the paper is organized as follows. Section \ref{part:nmf} details the technical aspects of NMF and describes the 3 approaches considered in this paper to achieve the task at hand. Section \ref{part:protocol} describes the environmental sound scene corpus and the experimental protocol setup. Section \ref{part:results} then describes and discusses the outcomes of the numerical experiments.

@@

Non-negative Matrix Factorization is a linear approximation method introduced by Lee and Seung, \cite{lee_learning_1999}, which can be used to approximate the spectrogram (obtained using a Short-Term Fourier Transform) of an audio file, $\mathbf{V}$, $\in \mathbb{R}^+_{F \times N}$ as :

@@

where $\mathbf{W} \in \mathbb{R}^+_{F \times K}$ is the \textit{dictionary} (or basis) matrix composed of audio spectrum and $\mathbf{H} \in \mathbb{R}^+_{K \times N}$ is the \textit{activation} matrix which summarizes the temporal evolution of each element of $\mathbf{W}$. An illustrative example can be found in Figure  \ref{fig:example_NMF}.

@@

and usually belongs to the $\beta-$divergence class \cite{fevotte_nonnegative_2009} in which the well known Euclidean distance (eq. \ref{eq:def_distEUC}), the Kullback-Leibler divergence (eq. \ref{eq:def_divKL}) and the Itakura-Saito divergence belong

@@

Algorithms have been proposed to solve the minimization problem (\ref{eq:min-D-WH}) iteratively such as the multiplicative update \cite{lee_algorithms_2000}, the alternating least square method \cite{cichocki_regularized_2007}, the projected gradient \cite{lin_projected_2007} \dots Here, the multiplicative update is chosen as it ensures convergence as proven in \cite{fevotte_algorithms_2011}.

\subsection{Supervised NMF}
First, supervised NMF (Sup-NMF) is used: the \textit{dictionary} includes audio spectrum of urban sound sources. In the urban environments, a lot of different sound sources present are known whose spectrum can be obtained and be a basis of $\mathbf{W}$. The \textit{activator} are then the unknown to estimate. In the first iteration, $\mathbf{H}$ is initialized randomly, then it is updated by the generic algorithm

@@

The supervised approach is useful when prior knowldge can be assumed for all the sources in the mixture, which is not a reasnable assumption in our application scenario. To some extend, prior knowledge can be considered for the traffic but not for the numerous kind of intereferences that can occur in a realistic scenario. To better take into account the diverse nature of urban scenes, semi-supervised NMF (Sem-NMF)\cite{lee_semi-supervised_2010} is good candidate as it offer more flexibility. This method assumes a \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of road traffic spectrum, and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated during optimization. In the literature, $J$ is set to a small number with respect to $K$ so as to force the optimization to still consider the fixed part of the dictionnary \cite{fevotte ?} \ml{mettre les parametres dnas la partie experimentale Here, $J = 2$}. The aim is to include in $\mathbf{W_r}$ the elements that are not related with the traffic. The problem (\ref{eq:nmf}) becomes

@@

Applications of Sem NMF for speech denoising from background noise or musical content can be found in \cite{joder2012real} and \cite{weninger2012supervised}.

@@

As it will be demonstrated in the experiments described in Section \ref{part:results}, those last approaches fails to provide consistent results in a wide range of urban areas and TIR ratios due generalization capabilities issues.

We therefore propose an alternative scheme based on the unsupervised NMF. Usually in unsupervised, $\mathbf{W}$, as  $\mathbf{H}$, is initialized randomly. Here, as the concerned sound source is known and audio samples of car passages are available, a initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain; see part \ref{part:dictionary_learning}. Then NMF is performed where $\mathbf{W}$ (eq. \ref{eq:updateW_unsup}) and $\mathbf{H}$ (eq.  \ref{eq:updateH_Sup}) are updated alternatively. $\mathbf{W}$ is therefore updated by forcing its initiation with \textit{a priori} knowledge but allowing it to adapt to the actual content of the scene under study.

@@

$D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right) = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ = 0 means that the elements are significantly different. This measure is bounded between 1 and 0 and is an invariant with respect to scale. \ml{ $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ is next sorted in descending order. } The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method. It is defined as:

@@

An illustrative example is shown on Figure \ref{fig:W_TI_NMF}.\\

@@

This approach is named \textit{Thresholded inititialized NMF} (TI NMF). Other thresholding methods as the \textit{soft} \cite{donoho1995noising} and the \textit{firm} \cite{fornasier2008iterative} and multiples way to process the similarity using a sigmoïd or a radial basis functions have been investigated. A fast parametric study revealed that the \textit{hard} thresholding method  with a linear representation of the similarity according to $K$, as in  Figure \ref{fig:W_TI_NMF}, was the most reliable approach.

@@

In order to validate the usefulness of the proposed NMF scheme to estimate the road traffic noise level, one need to compare the traffic noise level predicted by the algorithm to a reference level. The latter can hardly be measured or even annotated from real life recordings. Thus,  we propose to consider simulated sound scenes to assess the performance of the proposed NMF. This offers a controlled framework to design at low cost a wide diversity of sound environments in which all the traffic component is known allowing the computation of the reference level. %Thus, the estimated road traffic sound levels with the method can be compared to the real ones, introduced within each simulated sound scene.

@@

The corpus is designed with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simScene}}. \textit{SimScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound scenes in a .wav format by sequencing and summing audio samples that come from an isolated sound database. The simulator have been succesfully considered for a wide range of experimental design for sound detection algorithm assessment \ml{voir ma page de publi, taslp avec gregoire, taslp avec emmanouil, taslp avec mesaros}.

This database is divided into two categories: $i)$ the \textit{event} category which are the brief sounds (from 1 to 20 seconds) that are considered as salient including 245 sound event samples divided in 19 sound classes (\textit{ringing bell, birds, sweeping broom, car horn, car passages, hammer, drill, coughing, barking dog, rolling suitcase, closing door, plane, siren, footstep, storm, street noise, train, tramway, truck and voice}) and $ii)$ the \textit{background} or \textit{texture} category that includes all the sounds that are of long duration and whose acoustic properties do not vary with respect to time. 154 sound samples that belong to this category are divided in 9 sound classes (\textit{birds, construction site noise, crowd, park, rain, children playing in schoolyard, constant traffic noise, ventilation, wind}). The sound class \textit{car passages} comes from 60 recordings of 2 cars (Renault Megane and Renault Scenic) made on the Ifsttar's runway on different speeds with multiple gear ratio. The other audio files have been found online (\textit{freesound.org}) and within the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}. Each sound class is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots to allow some diversity in the resulting mixture).
The software allows the user to control some high level parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may bring some random behavior between the scenes. Furthermore, an audio file of each sound class present in the scene can be generated that allows to know its exact contribution as well as a text file that summarizes the time presence of all the events.\\

\begin{figure}[t]
    \centering
    \begin{subfigure}[t]{0.47\linewidth}
        \centering
       \includegraphics[width=\linewidth]{./figures/exempleSimScene3-spectrum.png}
       \caption{}
       \label{fig:simScene_spec}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.47\linewidth}
        \centering
       \includegraphics[width=\linewidth]{./figures/exempleSimScene3-timeDomain.png}
       \caption{}
       \label{fig:simScene_time}
    \end{subfigure}
    \caption{Example of a simple scene created with \textit{SimScene} software ((\ref{fig:simScene_spec}) the spectrogram, (\ref{fig:simScene_time}), the time domain}) with a sound background (road traffic in purple) and 3 sound events (car horn in red, car passage in green and whistling bird in blue):
    \label{fig:example_simScene}
\end{figure}

This database allows us to create a wide diversity of realistic urban sound scenes from the road traffic point of view \cite{gloaguen_creation_2017}. A sound mixing corpus is composed of 6 sub-corpus of 25 audio files each lasting 30 seconds. Each sub-corpus is characterized by a specific generic sound class that summed with traffic will make the estimation of the traffic level more difficult. The classes are: \textit{alert} (car horn, siren), \textit{animals} (barking dog , whistling birds), \textit{climate} (wind, rain), \textit{humans} (crowd noise and voice), \textit{mechanics} (different metallic and construction site noises) and \textit{transportation} (train, tramway and plane). In each file, traffic component is present as the sum of the background and event traffic sounds and is mixed with the other sound classes. The sound classes that are not related to the traffic component are summed up as the \textit{interfering} sound class. To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound scene, $L_{p,traffic}$, fixed to a specific level according to the sound level of the \textit{interfering} class, $L_{p,interfering}$,  following the relation (\ref{eq:tir}).

@@

The second estimator is the proposed scheme, based on the three NMF schemes presented in Section \ref{part:NMF}. Multiples experimental factors are involved here between the dictionary learning and NMF (see Figure \ref{fig:bloc_nmf}).

\begin{figure}

>>>>

\documentclass[a4paper,10pt]{article}

@@

The advent of low cost acoustic monitoring devices raises new interesting approaches for improving the monitoring of the acoustic quality of urban areas. State of the art approaches target road traffic noise maps and consider, as input, an estimate of the number and the speed of vehicles in major traffic lanes. Follows a prediction procedure that outputs an acoustic pressure level at any location in the modeled area.

Considering as input the acoustic pressure measured in many locations using a sensor grid approach would greatly complement and improve the quality of the predicted pressure values. Among the technical issues that raise this kind of innovative approaches, there is a need to identify which part of the overall acoustic pressure level is due to the road traffic.

In this paper, several techniques based on non-negative factorization framework are studied in this application scenario on a simulated sound scene corpus. The task being to the best of our knowledge never been considered in the literature, we propose an experimental protocol to validate the studied approaches that complies with standard reproducible research recommendations.

@@

With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps. These maps depict the sound level distribution over the city and an estimation of the number of city dwellers exposed to high noise levels. These maps play both an important communication role and help drawing up action plans to reduce noise exposure. Road traffic noise maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation modelling and express as output $L_ {DEN}$ and $L_N$ values, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively. Altough very useful, the produced noise maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015} or by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014}, despite the long data collection and calculation times. In addition, the usual road traffic noise maps are static, aggregating the exposure into indicators $L_{DEN}$ and $L_N$, that ignore the sound levels evolution throughout the day. 
The use of acoustic measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}. These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would make available of the long-term evolution of the traffic noise levels. It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time periods.

Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters ...), including the assessment of urban noise levels. The DYNAMAP project \cite{dynamap_2016} studied the deployment and feasibility of such installations focusing on sensor installations on specific roads at the city scale in Milan and Rome \cite{bellucci_life_2017}. 
The SONYC project (Sounds Of New-York City) aims to deploy a sensor network in New-York City for the purpose of monitoring constantly the noise pollution in the city \cite{mydlarz2017implementation}. In order to better know the urban sound environment, sensors are coupled with a detection tool that identifies the sound sources present \cite{salamon2017deep}. In a similar way, but reduced to few neighborhoods with a denser network, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{picaut2017characterization} aims to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, through data assimilation techniques.

Prior to data assimiliation, the issue of the correct estimation of the traffic sound level from acoustic measurements is still unsolved \cite{Mioduszewski}. Mainly because the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes \dots) that overlap. Consequently, the traffic sound level estimation based on measurements is not a trivial task.
Many recent works have focused on the detection or recognition tasks of environmental sounds \cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}. A two-step process is generally followed : describe the audio files with a set of features (Spectrum Gravity Spectrum, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks). A description of these features and classifiers can be found in \cite{cowling_comparison_2003} and their applications can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.

The main issue in the detection or recognition tasks is the overlap of environmental sounds. Although near major roads, traffic is predominant, there are many places where it overlaps with other sound sources which contribute significantly to the overall sound levels. To circumvent this issue, Socoro et al. propose to suppress time frames where there is significant overlap by considering an Anomalous Noise Events Detector \cite{socoro_anomalous_2017} which consist in detecting the unwanted sound sources from labeled recordings, \textit{i.e.} that are not related to the traffic component. Those time frames are then discarded in order not to take them into account during the estimation of the traffic sound level. 
An alternative approach that we will follow in this paper is to consider the blind source separation paradigm to reliably estimate the traffic noise level, see Figure \ref{fig:diagram}. That is, separating the contribution of the traffic from the other sources within a polyphonic scene. One major advantage of following such approach is that the estimate is continuously available, making the approach applicable in a wide range of urban areas, even where the traffic noise is relatively low compared to the remaining contributions.\\

In an urban environment context, source separation can be achieved with the help of acoustic microphone arrays and beamforming \cite{saruwatari2003blind}. However, this approach requires spreading multiple microphones arrays in cities that is very expensive (even with low cost microphones) and time-consuming for calibration and maintenance. This method is then not considered here to be deployed all over cities. In the opposite, monophonic sensor networks need less microphones but the main challenge is to succeed to estimate correctly the road traffic from only one signal in which all kind of sound sources can be present. A convenient method for this is the  Non-negative Matrix Factorization (NMF) technique \cite{lee_learning_1999}. When considering audio as input, it usually consists in approximating the magnitude spectrogram of an audio file by the product of two low rank matrices, one representing the components of interest and the other the contribution at a given time of those components to approximate the input magnitude spectrogram \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}. In the audio processing domain, NMF has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}. By design, this method deals reasonably well with the overlapping sound sources as soon as the overlap can be resolved on the time/frequency plane. 
Closer to our application scenario, NMF has been considered by Innami and Kasai \cite{satoshi_innami_nmf-based_2012}. After having performed NMF on simulated audio files, they perform a source separation in two steps by 1) separating the sound background from the events and 2) by isolating the events using spectral features using a $k$-means procedure.
We study in this paper different flavors of Non-Negative Matrix Factorization where traffic component is considered in its entirely whether it is a sound background or event. We demonstrate that supervised NMF and semi supervised NMF approaches have some interests but fail to give satisfactory results for the application at hand. We thus introduce another NMF scheme called thresholded initialized NMF that makes good use of prior knowledge about the source of interest, in our case the traffic noise, but also generalizes well to several kinds of urban areas and to traffic to interference ratio (TIR).

To perform the numerical experiments, we consider a corpus of simulated sound scenes created from a built-up sound database composed of a high number of diverse sound samples. The use of simulated sound scenes is mandatory for rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$), which would be difficult to extract from a recording of an urban scene.

\begin{figure*}[t]
\centering
\includegraphics[width=0.7\textwidth]{figures/bloc_diagram_source_separation.pdf}
\caption{Block diagram of the general source separation model}
\label{fig:diagram}
\end{figure*}

The remaining of the paper is organized as follows. Section \ref{part:nmf} details the technical aspects of NMF and describes the 3 approaches considered in this paper to achieve the task at hand. Section \ref{part:protocol} describes the corpus of environmental sound scenes and the experimental protocol setup. Section \ref{part:results} then describes and discusses the outcomes of the numerical experiments.

@@

Non-negative Matrix Factorization is a linear approximation method introduced by Lee and Seung, \cite{lee_learning_1999}, which can be used to approximate the spectrogram $\mathbf{\tilde{V}}$ (obtained using a Short-Term Fourier Transform) of an audio file, $\mathbf{V}$, $\in \mathbb{R}^+_{F \times N}$ as :

@@

where $\mathbf{W} \in \mathbb{R}^+_{F \times K}$ is the \textit{dictionary} (or basis) matrix composed of audio spectra and $\mathbf{H} \in \mathbb{R}^+_{K \times N}$ is the \textit{activation} matrix which summarizes the temporal evolution of each element of $\mathbf{W}$. An illustrative example can be found in Figure  \ref{fig:example_NMF}.

@@

and usually belongs to the $\beta-$divergence class \cite{fevotte_nonnegative_2009} in which the well known Euclidean distance (eq. \ref{eq:def_distEUC}) and the Kullback-Leibler divergence (eq. \ref{eq:def_divKL}) belong

@@

Algorithms have been proposed to solve the minimization problem (\ref{eq:min-D-WH}) iteratively such as the multiplicative update \cite{lee_algorithms_2000}, the alternating least square method \cite{cichocki_regularized_2007} or the projected gradient \cite{lin_projected_2007}. Here, the multiplicative update is chosen as it ensures convergence as proven in \cite{fevotte_algorithms_2011}.

\subsection{Supervised NMF}
First, supervised NMF (Sup-NMF) is used: the \textit{dictionary} includes audio spectra of urban sound sources. A lot of the different sound sources present in the urban environment are known whose their spectra can be obtained and be a basis of $\mathbf{W}$. The \textit{activation} matrix is then the unknown variable to estimate. In the first iteration, $\mathbf{H}$ is initialized randomly, then it is updated by the generic algorithm

@@

The supervised approach is useful when prior knowldge can be assumed for all the sources in the mixture, which is not a reasonable assumption in our application scenario. To some extend, prior knowledge can be considered for the traffic but not for the numerous kind of interferences that can occur in a realistic scenario. To better take into account the diverse nature of urban scenes, semi-supervised NMF (Sem-NMF)\cite{lee_semi-supervised_2010} is good candidate as it offers more flexibility. This method assumes a \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of road traffic spectra, and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated during optimization. In the literature, $J$ is set to a small number with respect to $K$ so as to force the optimization to still consider the fixed part of the dictionary %\cite{fevotte ?}
. The aim is to include in $\mathbf{W_r}$ the elements that are not related with the traffic. The problem (\ref{eq:nmf}) becomes

@@

Applications of Sem-NMF for speech denoising from background noise or musical content can be found in \cite{joder2012real} and \cite{weninger2012supervised}.

@@

As it will be demonstrated in the experiments described in Section \ref{part:results}, those last approaches fail to provide consistent results in a wide range of urban areas and TIR ratios due generalization capabilities issues.

We therefore propose an alternative scheme based on the unsupervised NMF. Usually in unsupervised, $\mathbf{W}$, as  $\mathbf{H}$, are initialized randomly. Here, as the concerned sound source is known and audio samples of car passages are available, an initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain; see part \ref{part:dictionary_learning}. Then NMF is performed where $\mathbf{W}$ (eq. \ref{eq:updateW_unsup}) and $\mathbf{H}$ (eq.  \ref{eq:updateH_Sup}) are updated alternatively. $\mathbf{W}$ is therefore updated by forcing its initiation with \textit{a priori} knowledge but allowing it to adapt to the actual content of the scene under study.

@@

$D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right) = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ = 0 means that the elements are completely different. This measure is bounded between 1 and 0 and is an invariant with respect to scale. $D_{\theta}\left(\mathbf{W_0} \vert \vert \mathbf{W} \right)$ is next sorted in descending order. The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method. It is defined as:

@@

An illustrative example is shown in Figure \ref{fig:W_TI_NMF}.\\

@@

This approach is named \textit{Thresholded inititialized NMF} (TI-NMF). Other thresholding methods as the \textit{soft} \cite{donoho1995noising} and the \textit{firm} \cite{fornasier2008iterative} and multiples way to process the similarity using a sigmoïd or a radial basis functions have been investigated. A fast parametric study revealed that the \textit{hard} thresholding method  with a linear representation of the similarity according to $K$, as in  Figure \ref{fig:W_TI_NMF}, was the most reliable approach.

@@

In order to validate the usefulness of the proposed NMF scheme to estimate the road traffic noise levels, one need to compare the traffic noise level predicted by the algorithm to a reference level. The latter can hardly be measured or even annotated from real life recordings. Thus,  we propose to consider simulated sound scenes to assess the performance of the proposed NMF. This offers a controlled framework to design at low cost a wide diversity of sound environments in which all the traffic component is known, thus allowing the computation of the reference level. %Thus, the estimated road traffic sound levels with the method can be compared to the real ones, introduced within each simulated sound scene.

@@

The corpus is designed with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simScene}}. \textit{SimScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound scenes in a .wav format by sequencing and summing audio samples that come from an isolated sound database. The simulator has been succesfully considered for a wide range of experimental design for sound detection algorithm assessment \cite{lafay:hal-01111381} \cite{benetos:hal-01520194} \cite{mesaros:hal-01650601}.

This database is divided into two categories: $i)$ the \textit{event} category, which are the brief sounds (from 1 to 20 seconds) that are considered as salient including 245 sound event samples divided in 19 sound classes (\textit{ringing bell, whistling bird, sweeping broom, car horn, passing car, hammer, drill, coughing, barking dog, rolling suitcase, closing door, plane, siren, footstep, storm, street noise, metallic noise, train, tramway, truck and voice}) and $ii)$ the \textit{background} or \textit{texture} category that includes all the sounds that are of long duration and whose acoustic properties do not vary with respect to time. 154 sound samples that belong to this category are divided in 9 sound classes (\textit{whistling bird, construction site noise, crowd noise, park, rain, children playing in schoolyard, constant traffic noise, ventilation, wind}). The sound class \textit{car passages} comes from 60 recordings of 2 cars (Renault Megane and Renault Scenic) made on the Ifsttar's runway at different speeds with multiple gear ratio. The other audio files have been found online (\textit{freesound.org}) and within the \textit{UrbanSound8k} database \cite{salamon_dataset_nodate}. Each sound class is composed of multiples samples (\textit{bird01.wav}, \textit{bird02.wav} \dots) to allow some diversity in the resulting mixture).
The software allows the user to control some high level parameters (number of events of each class that appear in the mixture, elapsed time between each sample of a same class, presence of a fade in and a fade out \dots) completed with a standard deviation that may bring some random behavior between the scenes. Furthermore, an audio file of each sound class present in the scene can be generated that allows to know its exact contribution as well as a text file that summarizes the time presence of all the events.\\

\begin{figure}[t]
    \centering
    \begin{subfigure}[t]{0.47\linewidth}
        \centering
       \includegraphics[width=\linewidth]{./figures/exempleSimScene3-spectrum.png}
       \caption{}
       \label{fig:simScene_spec}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.47\linewidth}
        \centering
       \includegraphics[width=\linewidth]{./figures/exempleSimScene3-timeDomain.png}
       \caption{}
       \label{fig:simScene_time}
    \end{subfigure}
    \caption{Example of a simple scene created with \textit{SimScene} software ((\ref{fig:simScene_spec}) the spectrogram, (\ref{fig:simScene_time}), the time domain) with a sound background (road traffic in purple) and 3 sound events (car horn in red, car passage in green and whistling bird in blue)}
    \label{fig:example_simScene}
\end{figure}

This database allows us to create a wide diversity of realistic urban sound scenes from the road traffic point of view \cite{gloaguen_creation_2017}. A sound mixing corpus is composed of 6 sub-corpus of 25 audio files each lasting 30 seconds. Each sub-corpus is characterized by a specific generic sound class that summed with traffic will make the estimation of the traffic level more difficult. The classes are: \textit{alert} (car horn, siren), \textit{animals} (barking dog , whistling bird), \textit{climate} (wind, rain), \textit{humans} (crowd noise and voice), \textit{mechanics} (metallic and construction site noises) and \textit{transportation} (train, tramway and plane). In each file, traffic component is present as the sum of the background and event traffic sounds and is mixed with the other sound classes. The sound classes that are not related to the traffic component are summed up as the \textit{interfering} sound class. To test different scenarios, each audio file is duplicated with the traffic sound level of the entire sound scene, $L_{p,traffic}$, fixed to a specific level according to the sound level of the \textit{interfering} class, $L_{p,interfering}$,  following the relation (\ref{eq:tir}).

@@

The second estimator is the proposed scheme, based on the three NMF schemes presented in Section \ref{part:nmf}. Multiples experimental factors are involved here between the dictionary learning and NMF (see Figure \ref{fig:bloc_nmf}), each experimental factor having multiples modalities.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/bloc_diagram_NMF_EN.pdf}
    \caption{Block diagram of the NMF experience with the dictionary design composed from a second sound database}
    \label{fig:bloc_nmf}
\end{figure}
